---
title: "Visium_pipeline"
output: html_document
date: "2023-09-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}

#removes all data from global environment
rm(list = ls())
```

## Import Libraries 
```{r librarys}
library(Seurat)
library(ggplot2)
```


## Setup Working Directories
```{r wd}
input_dir <- "I:/GMLSC1-Q2051/Pfizer/Visium/RAW_DATA/Pfizer/"
input_dir_sc <- "I:/GML001-Q1851/Andrew_C/Pfizer/scRNA/"
output_dir <- "I:/GML001-Q1851/Andrew_C/Pfizer/Visium/"
```


## Import Visium Data and run QC + Processing 
```{r Visium_Data_Inport}

###Load scRNA data list

lbtr_list <- readRDS(file = paste0(input_dir_sc, "scRNA_list.RDS"))


###Process Visium Data
import_visium <- function(sample, input_dir , output_dir, min_genes = 10, min_spots = 3, dim = 30){
  
  ### Set up data Directory
  outdir <- paste0(output_dir,sample,"/")
  dir_create(outdir)
  
  print(paste0("Import - Importing ", sample, " data..."))
  ###Load Visium data
  data <- Load10X_Spatial(data.dir = paste0(input_dir,sample, "/outs/"))
  
  
  ###Run QC
  
  # QC - removing Mitochondrial and Ribosomal Genes
  print("QC - removing Mitochondrial and Ribosomal Genes...")
  data <- PercentageFeatureSet(data, "^MT-", col.name = "percent_mito", assay="Spatial")
  data <- PercentageFeatureSet(data, "^RP[SL]", col.name = "percent_ribo", assay="Spatial")

  ## check the feature and count distribution before qc
  feats <- c("nFeature_Spatial", "nCount_Spatial")
        
  # filtering 
  selected_c <- WhichCells(data, expression = nFeature_Spatial > min_genes)
  length(selected_c)
    
  selected_f <- rownames(data)[Matrix::rowSums(data) > min_spots]
  length(selected_f)
    
  data <- subset(data, features = selected_f, cells = selected_c)
  
  print("Processing - Normalising data and running PCA...") 
  #Run Processing 
  d <- SCTransform(data, assay = "Spatial", verbose = FALSE)
  d <- RunPCA(d, assay = "SCT", verbose = FALSE)
  d <- FindNeighbors(d, reduction = "pca", dims = 1:dim)
  
  #save Normalised Object
  saveRDS(d, file = paste0(outdir,sample,"_normalised_seurat.RDS"))
  
  return(d)
}


#List of sample names
vis_sample_list <- c("VLP78_A",  "VLP78_D",  "VLP79_A",  "VLP79_D",  "VLP80_A",  "VLP80_D",  "VLP81_A",  "VLP82_A",  "VLP82_D",  "VLP83_A",  "VLP83_D")


#Inport and Normalise each sample
vis_data <- list()

for (sample in vis_sample_list){
  data <- import_visium(sample, input_dir = input_dir, output_dir = output_dir)
  vis_data <- c(vis_data, data)
  
}

saveRDS(vis_data, file = paste0(output_dir,"all_visium_normalised.RDS"))

```


# Run Visium Clustering
```{r vis_norm}

get_clusters <- function(data, res = 0.5, dim = 30){
  
  data <- FindClusters(data, resolution = res,verbose = FALSE)
  data <- RunUMAP(data, reduction = "pca", dims = 1:dim)
  
  return(data)
}


### Visium clusters with 0.2 res
vis_clusters <-list()

for(sample in vis_data[]){
  data <- get_clusters(sample, res = 0.2)
  vis_clusters <- c(vis_clusters,data)
}


```



## Plot visium clusters
```{r plot_vis_clust}


plot_visium_clusters <- function(data, sample, outdir, adding_graph = FALSE){
  
  clust_dir <- paste0(outdir,sample,"/clusters/")
  dir.create(clust_dir)
  
  plot <- SpatialDimPlot(data, group.by = "seurat_clusters")
  ggsave(paste0(clust_dir,sample,"_clusters.pdf"), plot)
  
  
  if (adding_graph == TRUE){
    #print("adding graph")
    return(plot)
  }
   
}

plot_list <- list()
## Plot data
idx <- 1
while (idx <= length(vis_sample_list)){
  plot <- plot_visium_clusters(vis_clusters[[idx]], vis_sample_list[idx], output_dir, i = idx)
  plot_list[[idx]] <- plot 
  idx <- idx + 1
}

```


## Run Label Transfer using MATCHED scRNA data
```{r label_transfer}


##using matching scRNA data

label_transfer <- function(sc_data, st_data, sample, outdir){
  
  lt_dir <- paste0(outdir,sample,"/label_transfer/")
  dir.create(lt_dir)
  
  print(paste0("Label Transfer - Running label transfer for ", sample, " matched scRNA and Visium Data ..."))
  anchors <- FindTransferAnchors(reference = sc_data, query = st_data, normalization.method = "SCT", verbose = FALSE)
  
  predictions.assay <- TransferData(anchorset = anchors, refdata = sc_data$celltype_major, prediction.assay = TRUE, weight.reduction = st_data[["pca"]], dims = 1:30 , verbose = FALSE)
  predictions.assay1 <- TransferData(anchorset = anchors, refdata = sc_data$celltype_major, prediction.assay = FALSE, weight.reduction = st_data[["pca"]], dims = 1:30 , verbose = FALSE)
  predictions.assay2 <- TransferData(anchorset = anchors, refdata = sc_data$celltype_minor, prediction.assay = FALSE, weight.reduction = st_data[["pca"]], dims = 1:30 , verbose = FALSE)
  predictions.assay3 <- TransferData(anchorset = anchors, refdata = sc_data$seurat_clusters, prediction.assay = FALSE, weight.reduction = st_data[["pca"]], dims = 1:30 , verbose = FALSE)
  
  #used for adding a pridction assay to seurat object
  st_data[["predictions_celltype_major"]] <- predictions.assay
 
  rownames(predictions.assay1) <- colnames(st_data)
  rownames(predictions.assay2) <- colnames(st_data)
  rownames(predictions.assay3) <- colnames(st_data)
  
  #Export Label transfer to file
  write.csv(predictions.assay1, file=paste0(lt_dir,sample,"_label_transfer.csv"))
  
  #Add top label for each spot to seurat object
  st_data$celltype_major <- predictions.assay1$predicted.id
  st_data$celltype_minor <- predictions.assay2$predicted.id
  st_data$celltype_clusters <- predictions.assay3$predicted.id
  
  saveRDS(st_data, file = paste0(lt_dir,sample,"_label_transfer.RDS"))
  
  return(st_data)
}


#Run data
vis_lb_tf <- list()

idx <-1 
while(idx <= length(vis_sample_list)){
  data <- label_transfer(lbtr_list[[idx]],vis_clusters[[idx]], vis_sample_list[idx], output_dir)
  vis_lb_tf <- c(vis_lb_tf, data)
  idx <- idx + 1
}

```


```{r}
#### Load Label Transfered Visium Objects
load_Visium <- function(sample, outdir){
  data <- readRDS(paste0(outdir,sample,"/label_transfer/",sample,"_label_transfer.RDS"))
  return(data)
}

vis_sample_list <- c("VLP78_A",  "VLP78_D",  "VLP79_A",  "VLP79_D",  "VLP80_A",  "VLP80_D",  "VLP81_A",  "VLP82_A",  "VLP82_D",  "VLP83_A",  "VLP83_D")
vis_lb_tf <- list()


for (sample in vis_sample_list){
  data <- load_Visium(sample,output_dir)
  vis_lb_tf <- c(vis_lb_tf, data)
}



```



## Plot Spatial Visualisation of Label Transfer 
```{r plot_label_transfer}



###Plots the proportions of each cell type
plot_lt <- function(data, sample, outdir, adding_graph = FALSE){
  
  lt_dir <- paste0(outdir,sample,"/label_transfer/")
  dir.create(lt_dir)
  
  DefaultAssay(data) <- "predictions_celltype_major"
  features <- rownames(data@assays$predictions_celltype_major@meta.features)
  plot <- SpatialFeaturePlot(data, features = features, ncol = 4)
  ggsave(paste0(lt_dir, sample,"_label_transfer_proportions.pdf"), plot, height = 15, width = 20)
  print(plot)
}


### Plots the major cell type label
cell_spots <- function(data, sample, outdir){
  
  lt_dir <- paste0(outdir,sample,"/label_transfer/")
  dir.create(lt_dir)
  
  plot1 <- SpatialDimPlot(data, group.by = "celltype_major")
  plot2 <- SpatialDimPlot(data, group.by = "celltype_minor")
  plots <- plot1+plot2
  ggsave(paste0(lt_dir, sample,"_cell_label_transfer.pdf"), plots)
  print(plots)
}



idx <- 1

while (idx <= length(vis_sample_list)){
  plot_lt(vis_lb_tf[[idx]], vis_sample_list[idx], output_dir)
  cell_spots(vis_lb_tf[[idx]], vis_sample_list[idx], output_dir)
  idx <- idx + 1
}


```



## Subset data to run Cell-Cell interaction analysis for tumour spots
```





```{r cluster_proportions_celltypes}

color_dict <- list(
  "Cancer Epithelial"="#A6CEE3",
  "Plasmablasts"= "#1F78B4",     
  "B-cells" = "#B2DF8A",      
  "T-cells" =  "#33A02C",     
  "CAFs" = "#FB9A99",          
  "Myeloid" =  "#E31A1C" ,        
  "Endothelial"= "#FDBF6F",
  "PVL"= "#FF7F00"                  
  )



celltype_proportion <- function(data, sample){
  cell_type_counts <- data@meta.data %>%
    group_by(seurat_clusters, label_transfer) %>%
    summarise(count = n())
  
 cell_type_counts <- cell_type_counts %>%
  mutate(color = unlist(color_dict[label_transfer]))
 
 unique_labels <- unique(cell_type_counts$label_transfer)
  
 # Create a named vector of corresponding colors
 legend_colors <- unlist(color_dict[unique_labels])

 plot <- ggplot(cell_type_counts, aes(y = count, x = seurat_clusters, fill = label_transfer, color = label_transfer)) +
    geom_bar(position = "fill", stat = "identity", width = 0.8) +
    scale_fill_manual(values = unlist(color_dict)) +
    scale_color_manual(values = unlist(color_dict)) +
    coord_flip() + theme_classic() + labs(x = "Cluster", y = "Proportion")
  
  
  ggsave(paste0(OUT_DIR,sample,"_celltype_proportions.pdf"))
  print(plot)
}

celltype_proportion(V82_A, "V82_A")
celltype_proportion(V82_D, "V82_D")
celltype_proportion(V83_A, "V83_A")
celltype_proportion(V83_D, "V83_D")



```



